---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# citoMove

The goal of citoMove is to provide an user-friendly R package for analyzing movement data with DNNs.

## Installation

The current version of citoMove depends on the **development** version of cito, which can be installed via:


```{r, eval = F}
# install.packages("devtools")
devtools::install_github('citoverse/cito')
```

You can install the development version of citoMove from [GitHub](https://github.com/) via:

```{r, eval = F}
devtools::install_github("citoverse/citoMove")
```

# Example - analysis of a deer movement track

```{r}
set.seed(123)
library(amt)
library(citoMove)
library(terra)
```

The following code creates a spatial covariate: the distance to the next forest.

```{r}
forest <- amt::get_sh_forest()
forest <- terra::subst(forest, 0, NA)
forest_dist <- distance(forest)
names(forest_dist) <- "forest_dist"
plot(forest_dist)
```

Next, we load the tracking data of one red deer individual. From amt documentation: 826 GPS relocations of one red deer from northern Germany. The data is already resampled to a regular time interval of 6 hours and the coordinate reference system is transformed to epsg:3035.

The follwoing code orders the data, creates random absences, and extract the environmental predictor (distance to forest) create before for all true and random steps. 

```{r}
dat_ssf <- amt::deer |> 
  steps_by_burst() |> 
  random_steps() |> 
  extract_covariates(forest_dist) |> 
  time_of_day() |> 
  mutate(case_ = as.integer(case_))
```

Now, we can fit the deep neural network (dnn) using citoMove: 

```{r}
model = dnn_ssf(case_ ~ forest_dist, data = dat_ssf, epoch = 30L, plot = T, verbose = FALSE)
```

The plot that is produced visualizes the improvement of the model during training. The default architecture of the neural networks entails two hidden layers. We can visualize the architecture via 

```{r, message=F}
plot(model)
```

A summary of the fitted model is provided via 

```{r}
summary(model)
```

The cito package contains several options to create effect plots. Our current recommendation is to use accumulated local effect (ALE) plots to visualize the estimated selection effects

```{r}
ALE(model)
```
